Measuring Individual Partisan Harm in Redistricting

### 2021-09-13 paper/paper.Rmd
long awaited section 5

@@

Over the past decade [**CM: correct?!**], researchers have developed a veritable zoo of measures to identify partisan gerrymanders.
Specifically, these measures aim to quantify deviations from partisan symmetry, fairness, or competitiveness.
Some of the most commonly used are perhaps the *efficiency gap* [@stephanopoulos2015], *partisan bias* [@tufte1973pb; @king1987democratic], and the *mean-median* measure [@mcdonald2015unfair; @wang2016three].
All of these measures may be estimated using district vote shares.

@@

1.  has been criticized [@tamcho2017egap; @bernstein2017egap; @deford2020symmetry]

## Partisan Bias

1.  Integrating yields a related measure of @nagle2015measures, which has also been criticized by @deford2020symmetry.
2.  Focuses on only one part of S-V curve, which may not be relevant

## Mean-Median

1.  Like PB, focuses on only one part of S-V curve, which may not be relevant

## Districting Plan Competitiveness {#sec:competitive}

@@

>>>>

<!--  Section 1 here should really be what we're interested in I think -->

@@

In the years since Vieth [@2004vieth] researchers have developed a veritable zoo of measures to identify partisan gerrymanders.
Specifically, these measures aim to quantify deviations from partisan symmetry, fairness, or competitiveness.
Some of the most commonly used are perhaps the *efficiency gap* [@stephanopoulos2015], *partisan bias* [@tufte1973pb; @king1987democratic], and the *mean-median* measure [@mcdonald2015unfair; @wang2016three].
All of these measures may be estimated using district vote shares.
We now discuss the relationship between our measure and these three measures.
<!-- chris thinks we shouldn't cite wang here because he copied mcdonald -->

@@

The efficiency gap interrogates how a party's votes were wasted in an election under a districting system.
By comparing the wasted votes for Democrats to the wasted votes for Republicans, rescaled by the total votes cast, it offers a signed percent.
Interpretation of the efficiency gap is intended to be straightforward and usable by courts.
The sign indicates that a plan favors one party over another, while the magnitude indicates the degree of preference.
The general recommendation for use in court is that a bias amounting to the size of two Congressional seats or 8% for state legislative plans would be deemed unacceptable.
When we unpack the idea behind the efficiency gap, we see that it shares significant intuition with our measure $F(q)$.
Both measures try to understand how individuals are rendered unable to effectively participate in democratic processes.

Various criticisms of the efficiency gap have been made due to its potential for volatility in noncompetitive elections [@tamcho2017egap; @bernstein2017egap; @deford2020symmetry].
By directly including simulated plans and electoral data in the formation of $F(q)$, despite shared intuition with the efficiency gap, our measure is not prey to this susceptibility.

```{=html}
<!--
Further literature indicates that while the efficiency gap was originally introduced as a measure of partisan symmetry, "it does not imply a coherent seats--votes curve" and thus cannot be a measure of partisan symmetry [@katz2020].
-->
```
## Partisan Bias

Partisan bias can be considered the classic measure of partisan symmetry.
The bias is the number of seats that would need to be reallocated to make the system symmetric at a chosen statewide proportion of votes, often 0.50.
It should not be surprising then that these measures differ often and greatly.

Further, by relying on alternative districting plans, we do not necessitate the use of partisan swings.
This is of great benefit in scenarios where swings must be so significant that they stray far from the data.
Instead, the measures of harm and utility rely on $votes^{p}$ which can be model-free estimates of electoral performance of a party (namely historical election returns) or can be some sort of model, such as precinct-level election forecasts.
Of course, the votes could even be generated by notions of partisan swings, but as a choice rather than a requirement.

Notably, as the quantity of interest is not partisan symmetry, our measures do not require the existence of a seats-votes curve.
This is consequential for the study of congressional redistricting.
As a seats-votes curve is defined for a legislature, rather than a delegation, to understand the impact of single state's congressional map, you would be forced to consider how every other state drew their maps [@katz:king:rose:21a].
Thus, if this taken as the definition of a seats-votes curve, then measures of partisan symmetry lack diagnostic power to identify gerrymanders for the delegations that academics, legal scholars, and the general public are often focused on.
The individualized focus of our measures are, however, portable to congressional, state legislative, or any other form of districting and elections.

```{=html}
<!--
1.  Integrating yields a related measure of @nagle2015measures, which has also been criticized by @deford2020symmetry.
2.  Focuses on only one part of S-V curve, which may not be relevant
-->
```
## Mean-Median Difference

Our measures greatly differ from the Mean-Median Difference.
With the focus on individuals, the considerations of the mean district are only related to individuals in-so-far as they describe the general propensity to be harmed or benefited by a particular map.
The consideration of a median district is similarly distanced.
As with partisan bias, as our measures do not attempt to measure partisan symmetry, our measures can be applied to congressional delegations.

```{=html}
<!--
1.  Like PB, focuses on only one part of S-V curve, which may not be relevant
-->
```
## Districting Plan Competitiveness {#sec:competitive}

@@

We should say something about the importance in developing further measures which utilize simulation approaches
